---
title: "06_TEconsensus_analysis.Rmd"
author: "Xinyu Xiang"
date: "2024-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyr)
library(dplyr)
library(stringr)
library(ggrastr)
library(DESeq2)
library(pheatmap)
library(ComplexHeatmap)


```

# 00. configure info
```{r}
metadata <- read.table("./metadata_merge.txt", stringsAsFactors = FALSE, header=TRUE)
metadata
piRNAlen.cutoff = 23

TE_config <- read.delim("/reference_AmexG_v6.0/AmexG_v6.0_TE_xinyu_anno/AmexG_v6.0_consensus.merge.config", stringsAsFactors = FALSE, header=FALSE, col.names = c("TEID", "class", "family","TE","length"))

TEclass_order <- c("LTR", "DNA", "LINE", "DIRS", "RC", "PLE", "SINE", "Ngaro", "Others")

TEclass_col <- c(LTR  = "#9b2226", DNA  = "#ca6702", LINE  = "#007f5f", SINE  = "#4895ef", DIRS  = "#e9d8a6", Ngaro  = "#b9fbc0", PLE  = "#e7bc91", RC = "#dbcdf0", Others  = "#eddcd2")
condition_col <- c(testis = "#354f52", oocyte = "#9f86c0")
type_col = c(testis_fwd = "#286566", oocyte_fwd = "#5c2862",
              testis_rev = "#3f9497", oocyte_rev = "#873388")



```

# 01. piRNA coverage track over TE consensus sequence

## 1.1 calculate total count
```{r}
####### calculate total count ####### 
count.sum.ls <- list()
for (Sample in metadata$sample) {
length_summary <- read.table(paste0("./03_library_info/",Sample,"_len.txt"), col.names = c("sRNA_length", "sRNA_count") )
count.sum.ls[[Sample]] <- c(countsum = sum(length_summary$sRNA_count))
}
count.sum.df <- data.frame(sample = metadata[,"sample"], count = do.call(rbind, count.sum.ls)) 
write.table(count.sum.df, "./11_views_TE_v2/01_TE_track/totalcount.txt", quote = F, row.names = F)

```


## 1.2 track for piRNA
plot by ggplot2
```{r}
count.sum.df = read.table("./11_views_TE_v2/01_TE_track/totalcount.txt", header = T)

for (j in 1:nrow(TE_config)) {
  TEclass <- TE_config[j,"class"]
  TEfamily <- gsub("/","-",TE_config[j,"family"])
  TEsubfamily <- TE_config[j,"TE"]

  piRNA.fwd.ls <- list()
  piRNA.rev.ls <- list()
  for (Sample in metadata$sample) {
    ##### input TEcov.bed #####
    ### fwd strand ###
    piRNA.fwd.ls[[Sample]] <- read.delim(paste0("./11_views_TE_v2/01_TE_track/TEcovbed_piRNA/", Sample, ".", TEsubfamily, ".fwd.bed"), stringsAsFactors = FALSE, header = FALSE, col.names = c("TEID", "start", "end", "count") ) %>% 
      mutate(sample = Sample, condition = sub("..$", "", Sample), 
             count = as.numeric(replace(count, count == "NAN", 0)))

    ### rev strand ###
    piRNA.rev.ls[[Sample]] <- read.delim(paste0("./11_views_TE_v2/01_TE_track/TEcovbed_piRNA/", Sample, ".", TEsubfamily, ".rev.bed"), stringsAsFactors = FALSE, header = FALSE, col.names = c("TEID", "start", "end", "count") ) %>% 
      mutate(sample = Sample, condition = sub("..$", "", Sample),
             count = as.numeric(replace(count, count == "NAN", 0)))
  }
  
    # merge for all samples and calculate RPM
    piRNA.fwd.RPM.df <- do.call(rbind, piRNA.fwd.ls) %>% left_join(count.sum.df) %>% 
      mutate(RPM = round(1000000 * count / countsum, 2), strand = "fwd") 
    piRNA.rev.RPM.df <- do.call(rbind, piRNA.rev.ls) %>% left_join(count.sum.df) %>% 
      mutate(RPM = round(-1000000 * count / countsum, 2), strand = "rev") 
    
    # calculate RPM_mean
    piRNA.RPM.df <- rbind(piRNA.fwd.RPM.df, piRNA.rev.RPM.df)
    piRNA.RPM.mean.df <- piRNA.RPM.df %>% 
      group_by(condition, strand, TEID, start, end) %>% summarise(RPM_mean = mean(RPM)) %>%
      mutate(type = paste0(condition, "_", strand),
             condition = factor(condition, levels = c("testis", "oocyte")))
    write.csv(piRNA.RPM.mean.df, paste0("./11_views_TE_v2/01_TE_track/TEtrackplot_piRNA_RPMdata/", TEclass, ".", TEfamily, ".", TEsubfamily, ".RPM_mean.csv"), quote = F, row.names = F)
    
    
    ## plot ##
    plot_width <- round(max(piRNA.RPM.mean.df$end) / 20, 0)
    pdf(paste0("./11_views_TE_v2/01_TE_track/TEtrackplot_piRNA/", TEclass, ".", TEfamily, ".", TEsubfamily, ".pdf"), 
        height = 5, width = plot_width)
       ymax <- max(abs(piRNA.RPM.mean.df$RPM_mean))
       
       p <- ggplot(piRNA.RPM.mean.df, aes(x = end, y = RPM_mean, fill = type)) +
         geom_col(position = "identity") +
         labs(x = "position", y = "RPM mean", title = paste0(TEclass, ".", TEfamily, ".", TEsubfamily)) + 
         scale_y_continuous(limits = c(-ymax, ymax)) +
           scale_x_continuous(
             breaks = seq(0, max(piRNA.RPM.mean.df$end), by = 100), 
             minor_breaks = seq(0, max(piRNA.RPM.mean.df$end), by = 10)  ) +
         scale_fill_manual(values = type_col) + 
         facet_grid( ~ condition) +
         theme_minimal() +
         theme(panel.grid.minor.y = element_blank())
       
       plot(p)
    dev.off()
 
}



```



# 02. barplot of top piRNA enrichment by TE family / piRNA RPKM bin by TE class
## 2.1 input & clean TE alignment bed
```{r}
TE_bed_stat_ls <- list()
for (Sample in metadata$sample) {
  print(Sample)
  TE_bed <- read.table(paste0("./04_STAR_TE_mismatch3_overlap15/",Sample,"_TE_srt_edit.bed"), header = FALSE, stringsAsFactors = FALSE, comment.char = "", col.names = c("TEID", "Start", "End", "readID", "length", "strand", "NH"))
  TE_bed_stat_ls[[Sample]] <- TE_bed %>% 
    filter(length >= piRNAlen.cutoff) %>% # filter piRNAs
    mutate(sRNA_count = round(1/ NH, 3)) %>% # sRNA count score for mapped site
    group_by(TEID) %>%
    summarise(count_sum = sum(sRNA_count)) %>%
    mutate(sample = Sample)
}

TE_bed_stat <- do.call(rbind, TE_bed_stat_ls) %>% left_join(TE_config)

```


## 2.2 calculate RPKM
```{r}
TE_countsum <- TE_bed_stat %>%
  group_by(sample) %>%
  summarise(total_count = sum(count_sum))

# by TE
TE_bed_RPKM <- TE_bed_stat %>% left_join(TE_countsum) %>%
  mutate(RPKM = round(count_sum *1000000 *1000/ (total_count * length), 3))
head(TE_bed_RPKM)

write.csv(TE_bed_RPKM, paste0("./11_views_TE_v2/01_sRNA_statistics/pirna_countRPKM_byTE.csv"), quote = F, row.names = F)

# by TE family
TEfamily_countsum <- TE_bed_stat %>%
  group_by(sample, family, class) %>%
  summarise(familycount_sum = sum(count_sum),
            length_sum = sum(length))


TEfamily_bed_RPKM <- TEfamily_countsum %>% left_join(TE_countsum) %>%
  mutate(RPKM = round(familycount_sum *1000000 *1000/ (total_count * length_sum), 3))
head(TEfamily_bed_RPKM)

write.csv(TEfamily_bed_RPKM, paste0("./11_views_TE_v2/01_sRNA_statistics/pirna_countRPKM_byTEfamily.csv"), quote = F, row.names = F)



```


## 2.3 barplot of top piRNA enrichment by TE family
```{r}
##### input data #####
TEfamily_bed_RPKM <- read.csv(paste0("./11_views_TE_v2/01_sRNA_statistics/pirna_countRPKM_byTEfamily.csv"))

##### plot barplot #####
pdf(paste0("./11_views_TE_v2/02_barplot_countRPKM/barplot_pirna_countRPKM_byTEfamily.pdf"), height = 5, width = 13)
##### plot by TE family, sample mean #####
plot.df <- TEfamily_bed_RPKM %>% group_by(family, class) %>%
  summarise(RPKMmean = mean(RPKM)) %>%
  mutate(log2RPKMmean = round(log2(RPKMmean +1), 3))

family_order <- plot.df %>% arrange(desc(RPKMmean)) %>% select(family) %>% distinct()

plot.df <- plot.df %>%
  mutate(family = factor(family, levels = family_order$family), 
        class = factor(class, levels = TEclass_order))
head(plot.df)

y_max <- ceiling(max(plot.df$log2RPKMmean))
p <- ggplot(plot.df, aes(x=family, y = log2RPKMmean, fill=class)) + 
      geom_col(position = "dodge", width = 0.8) + #geom_col(position = position_dodge2(width = 0.5, preserve = "single")
      labs(x = "TE family", y = "log2(RPKMmean+1)", title = paste0("TE consensi count RPKMmean")) + 
      scale_fill_manual(values = TEclass_col) + 
      facet_grid( ~ class, scales = "free_x", space = "free_x") + # group by class #
      scale_y_continuous(limits = c(0, y_max)) +
      theme_classic() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1)) # text = element_text(family = "Arial"), # ; legend(x = "topright", legend = names(col1), fill = col1)
  
  plot(p)

  
##### plot by TE family, by each sample #####
family_order <- TEfamily_bed_RPKM %>% arrange(desc(RPKM)) %>% select(family) %>% distinct()

plot.df <- TEfamily_bed_RPKM %>%
  mutate(log2RPKM = round(log2(RPKM +1), 3),
    family = factor(family, levels = family_order$family), 
    class = factor(class, levels = TEclass_order))
head(plot.df)

p <- ggplot(plot.df, aes(x=family, y = log2RPKM, fill=class)) + 
      geom_col(position = "dodge", width = 0.8) + #geom_col(position = position_dodge2(width = 0.5, preserve = "single")
      labs(x = "TE family", y = "log2(RPKM+1)", title = paste0("TE consensi count RPKM")) + 
      scale_fill_manual(values = TEclass_col) + 
      facet_grid(sample ~ class, scales = "free_x", space = "free_x") + # group by class #
      scale_y_continuous(limits = c(0, y_max)) +
      theme_classic() +
      theme(axis.text.x = element_text(angle = 60, hjust = 1)) # text = element_text(family = "Arial"), # ; legend(x = "topright", legend = names(col1), fill = col1)
  
  plot(p)
  
  
##### plot by TE family, testis/oocye mean #####
# testis
 TEfamily_bed_RPKM_bytestis <- TEfamily_bed_RPKM %>% 
    filter(sample %in% c("testis-1", "testis-2")) %>%
    group_by(family, class) %>%
    summarise(RPKM_mean = mean(RPKM))
  family_order <- TEfamily_bed_RPKM_bytestis %>% arrange(desc(RPKM_mean)) %>% select(family) %>% distinct()
  
  plot.df <- TEfamily_bed_RPKM_bytestis %>%
    mutate(log2RPKM_mean = round(log2(RPKM_mean +1), 3),
      family = factor(family, levels = family_order$family), 
      class = factor(class, levels = TEclass_order))
  head(plot.df)
  
  p1 <- ggplot(plot.df, aes(x=family, y = log2RPKM_mean, fill=class)) + 
        geom_col(position = "dodge", width = 0.8) + #geom_col(position = position_dodge2(width = 0.5, preserve = "single")
        labs(x = "TE family", y = "log2(RPKMmean+1)", title = paste0("TE consensi count RPKM, testis")) + 
        scale_fill_manual(values = TEclass_col) + 
        facet_grid( ~ class, scales = "free_x", space = "free_x") + # group by class #
        scale_y_continuous(limits = c(0, y_max)) +
        theme_classic() +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) # text = element_text(family = "Arial"), # ; legend(x = "topright", legend = names(col1), fill = col1)
    plot(p1)

# oocyte
 TEfamily_bed_RPKM_byoocyte <- TEfamily_bed_RPKM %>% 
    filter(sample %in% c("oocyte-1", "oocyte-2")) %>%
    group_by(family, class) %>%
    summarise(RPKM_mean = mean(RPKM))
  family_order <- TEfamily_bed_RPKM_byoocyte %>% arrange(desc(RPKM_mean)) %>% select(family) %>% distinct()
  
  plot.df <- TEfamily_bed_RPKM_byoocyte %>%
    mutate(log2RPKM_mean = round(log2(RPKM_mean +1), 3),
      family = factor(family, levels = family_order$family), 
      class = factor(class, levels = TEclass_order))
  head(plot.df)
  
  p2 <- ggplot(plot.df, aes(x=family, y = log2RPKM_mean, fill=class)) + 
        geom_col(position = "dodge", width = 0.8) + #geom_col(position = position_dodge2(width = 0.5, preserve = "single")
        labs(x = "TE family", y = "log2(RPKMmean+1)", title = paste0("TE consensi count RPKM, oocyte")) + 
        scale_fill_manual(values = TEclass_col) + 
        facet_grid( ~ class, scales = "free_x", space = "free_x") + # group by class #
        scale_y_continuous(limits = c(0, y_max)) +
        theme_classic() +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) # text = element_text(family = "Arial"), # ; legend(x = "topright", legend = names(col1), fill = col1)
    plot(p2)
  
##### plot by TE family, sample separate #####
for (Sample in metadata$sample) {
 TEfamily_bed_RPKM_bysample <- TEfamily_bed_RPKM %>% filter(sample == Sample)
  family_order <- TEfamily_bed_RPKM_bysample %>% arrange(desc(RPKM)) %>% select(family) %>% distinct()
  
  plot.df <- TEfamily_bed_RPKM_bysample %>%
    mutate(log2RPKM = round(log2(RPKM +1), 3),
      family = factor(family, levels = family_order$family), 
      class = factor(class, levels = TEclass_order))
  head(plot.df)
  
  p <- ggplot(plot.df, aes(x=family, y = log2RPKM, fill=class)) + 
        geom_col(position = "dodge", width = 0.8) + #geom_col(position = position_dodge2(width = 0.5, preserve = "single")
        labs(x = "TE family", y = "log2(RPKM+1)", title = paste0("TE consensi count RPKM, ", Sample)) + 
        scale_fill_manual(values = TEclass_col) + 
        facet_grid( ~ class, scales = "free_x", space = "free_x") + # group by class #
        scale_y_continuous(limits = c(0, y_max)) +
        theme_classic() +
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) # text = element_text(family = "Arial"), # ; legend(x = "topright", legend = names(col1), fill = col1)
    plot(p)
}  
  
dev.off()




```



# 05. five end distence
## 5.1 line chart of five end distence freq
```{r}
merge.ls <- list()
plot.ls <- list()
for (j in 1:nrow(TE_config)) {
  TEclass <- TE_config[j,"class"]
  TEfamily <- gsub("/","-",TE_config[j,"family"])
  TEsubfamily <- TE_config[j,"TE"]
  
  gap_count.ls <- list()
  for (Sample in metadata$sample) {
    input_file <- paste0("./11_views_TE_v2/06_5enddis/",Sample,".",TEsubfamily,".gap_count.txt")
    ### input and calculate frequency ###
    if (file.info(input_file)$size > 0) {
    gap_count.ls[[paste(TEsubfamily, Sample)]] <- read.table(input_file, sep="\t", stringsAsFactors = FALSE, header = FALSE)
    colnames(gap_count.ls[[paste(TEsubfamily, Sample)]]) <- c("gap", "count")
    gap = data.frame(gap = c(0:30))
    gap_count.ls[[paste(TEsubfamily, Sample)]] <- gap %>% 
      left_join(gap_count.ls[[paste(TEsubfamily, Sample)]]) %>% replace(is.na(.), 0) %>% 
      mutate(count_sum = sum(count)) %>% mutate(freq = round(count/count_sum ,3)) %>% 
      mutate(sample = Sample, condition = sub("..$", "", Sample)) %>%
      select("gap", "freq", "sample", "condition")
    
    } else {
      gap_count.ls[[paste(TEsubfamily, Sample)]] <- data.frame(gap = c(0:30), 
                                                               freq = rep(0,31), 
                                                               sample = Sample, 
                                                               condition = sub("..$", "", Sample) )}
    
  }
  
  # merge for all samples
  merge.ls[[TEsubfamily]] <- do.call(rbind, gap_count.ls)
  # calculate mean and sd for freq
  plot.ls[[TEsubfamily]] <- merge.ls[[TEsubfamily]] %>% 
    group_by(condition, gap) %>% 
    summarize(mean_freq = mean(freq), sd_freq = round(sd(freq),3)) %>% 
    mutate(TE = TEsubfamily, family = TEfamily, class = TEclass)

  ### plot frequency of piRNA sense-antisense 5'end distance ###
  pdf(paste0("./11_views_TE_v2/06_5enddis_linechart/", TEclass, ".", TEfamily, ".", TEsubfamily, ".pdf"), height = 4, width = 7)
  p1 <- ggplot(data=plot.ls[[TEsubfamily]], aes(x=gap, y=mean_freq, fill = condition, colour = condition)) + 
    geom_line(linewidth = 0.5, stat="identity") + 
    geom_errorbar(aes(ymin = mean_freq - sd_freq, ymax = mean_freq + sd_freq), width = 0.2) +
    scale_colour_manual(values = condition_col) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0), limits = c(0, 30)) +
    theme_classic(base_size = 15) +
    labs(title=paste0("piRNA 5' end dis, ",TEclass,".",TEfamily,".",TEsubfamily), x="Nucleotides", y="Frequency") 
  plot(p1)
  dev.off()

}

plot.df <- do.call(rbind, plot.ls)
write.csv(plot.df, paste0("./11_views_TE_v2/06_5enddis_linechart/00_5enddis_freq.csv"), quote = F, row.names = F)



```

## 5.2 heatmap of five end distence freq
```{r}
data.df <- read.csv(paste0("./11_views_TE_v2/06_5enddis_linechart/00_5enddis_freq.csv"))
head(data.df)

## row annotation
row_annotation_df <- data.df %>%
  select(TE, class) %>%
  distinct() %>% # keep the original arranged order
  as.data.frame() %>%
  mutate(class = factor(class, levels = TEclass_order) ) %>%
  arrange(class, TE) %>% # reorder
  tibble::column_to_rownames(var = "TE")
         
head(row_annotation_df)

row_annotation <- HeatmapAnnotation(df = row_annotation_df,
                                    col = list(class = TEclass_col),
                                    which = "row")

## build data matrix for plot
testis.matrix <- data.df %>% filter(condition == "testis") %>%
  select(TE, gap, mean_freq) %>%
  spread(key = gap, value = mean_freq, fill = NA) %>%
  as.data.frame() %>%
  tibble::column_to_rownames(var = "TE")
# reorder by row_annotation
testis.matrix <- testis.matrix[rownames(row_annotation_df),]

oocyte.matrix <- data.df %>% filter(condition == "oocyte") %>%
  select(TE, gap, mean_freq) %>%
  spread(key = gap, value = mean_freq, fill = NA) %>%
  as.data.frame() %>%
  tibble::column_to_rownames(var = "TE")
# reorder by row_annotation
oocyte.matrix <- oocyte.matrix[rownames(row_annotation_df),]

write.csv(testis.matrix, paste0("./11_views_TE_v2/06_5enddis_heatmap/HeatmapOf5enddis.mean_freq.testis.csv"), quote = F, row.names = F)
write.csv(oocyte.matrix, paste0("./11_views_TE_v2/06_5enddis_heatmap/HeatmapOf5enddis.mean_freq.oocyte.csv"), quote = F, row.names = F)



  
### plot heatmap ###
col1 <- colorRamp2(c(0, 0.1, 0.45), c("#023e8a", "white", "#9e2a2b"))
pdf(file=paste0("./11_views_TE_v2/06_5enddis_heatmap/HeatmapOf5enddis.pdf"),height=7,width=7)
### plot all ###
Heatmap(testis.matrix, 
        column_title = "mean of 5enddis freq, testis",
        left_annotation = row_annotation,
        cluster_columns = FALSE,
        cluster_rows = FALSE, row_dend_reorder = FALSE, # keep original order
        show_row_names = FALSE, show_column_names = TRUE,
        col = col1,
        # row_split = row_split, # split by row
        name = "mean freq",
        use_raster = FALSE # keep original resolution
        )

Heatmap(oocyte.matrix, 
        column_title = "mean of 5enddis freq, oocyte",
        left_annotation = row_annotation,
        cluster_columns = FALSE,
        cluster_rows = FALSE, row_dend_reorder = FALSE, # keep original order
        show_row_names = FALSE, show_column_names = TRUE,
        col = col1,
        # row_split = row_split, # split by row
        name = "mean freq",
        use_raster = FALSE # keep original resolution
        )
  dev.off()
  
```


# 06. 1U10A
## 6.1 barplot
```{r}
merge.ls <- list()
plot.ls <- list()
for (j in 1:nrow(TE_config)) {
  TEclass <- TE_config[j,"class"]
  TEfamily <- gsub("/","-",TE_config[j,"family"])
  TEsubfamily <- TE_config[j,"TE"]

  nuc_count.ls <- list()
  Nucleotide.df = data.frame(position = c(rep("First",4),rep("Tenth",4)), nucleotide = rep(c("A","T","C","G"),2))
  for (Sample in metadata$sample) {
    input_file <- paste0("./11_views_TE_v2/07_1U10A/piRNA.",Sample,".",TEsubfamily,".count_nuc.txt")
    file.size <- file.info(input_file)$size
    file.size[is.na(file.size)] <- 0
    
    ### input and calculate frequency ###
    if (file.size > 0) {
    nuc_count.ls[[paste(TEsubfamily, Sample)]] <- read.table(input_file, sep="\t", stringsAsFactors = FALSE, header = FALSE, 
                                    col.names = c("position", "nucleotide", "count")) %>% 
      mutate(nucleotide = as.character(nucleotide))

    nuc_count.ls[[paste(TEsubfamily, Sample)]] <- Nucleotide.df %>% 
      left_join(nuc_count.ls[[paste(TEsubfamily, Sample)]]) %>% 
      replace(is.na(.), 0) %>% 
      group_by(position) %>% 
      mutate(count_sum = sum(count)) %>% mutate(freq = (count/count_sum)) %>% 
      mutate(sample = Sample, condition = sub("..$", "", Sample)) %>% 
      select("position", "nucleotide", "freq", "sample", "condition")

    
    } else {nuc_count.ls[[paste(TEsubfamily, Sample)]] <- data.frame(position = c(rep("First",4),rep("Tenth",4)), 
                                                                     nucleotide = rep(c("A","T","C","G"),2), 
                                                                     freq = rep(0,8), 
                                                                     sample = Sample, 
                                                                     condition = sub("..$", "", Sample) )}
  }
  
  # merge for all samples
  merge.ls[[TEsubfamily]] <- do.call(rbind, nuc_count.ls)
  # calculate mean and sd for freq
  plot.ls[[TEsubfamily]] <- merge.ls[[TEsubfamily]] %>% 
    mutate(condition = factor(condition, levels = c("testis", "oocyte")),
           nucleotide = factor(if_else(nucleotide == "T", "U", nucleotide), levels = c("A", "U", "C", "G"))) %>%
    group_by(condition, position, nucleotide) %>% 
    summarize(mean_freq = mean(freq), sd_freq = round(sd(freq),3)) %>% 
    mutate(TE = TEsubfamily, family = TEfamily, class = TEclass) 
  
  nucleotide_col = c(A = "#3ec300", U = "#e13700", C = "#118ab2", G = "#2d3047" )
  pdf(paste0("./11_views_TE_v2/07_1U10A_barplot/",TEclass,".",TEfamily,".",TEsubfamily,".pdf"), height = 4, width = 4)
  p1 <- ggplot(data=plot.ls[[TEsubfamily]], aes(x=position, y=mean_freq, fill=nucleotide)) + 
    geom_bar(colour = "black", stat = "identity", position = "fill", width = 0.8) + 
    theme_classic(base_size = 12) + scale_fill_manual(values = nucleotide_col) +
    # theme(axis.text.x = element_text(angle = 90)) + 
    # scale_y_continuous(expand = c(0,0)) +
    facet_grid( ~ condition) + #, scales = "free_x", space = "free_x"
    labs(x = "Nucleotide", y = "mean proportion", title = "mapped piRNA composition")
  plot(p1)
  
  dev.off()
}


plot.df <- do.call(rbind, plot.ls) %>% mutate(mean_freq = round(mean_freq, 3))
write.csv(plot.df, paste0("./11_views_TE_v2/07_1U10A_barplot/00_1U10A_freq.csv"), quote = F, row.names = F)

```

## 6.2 heatmap
```{r}
data.df <- read.csv(paste0("./11_views_TE_v2/07_1U10A_barplot/00_1U10A_freq.csv")) %>% 
  mutate(Rowname = paste0(condition, "-", position, "-", nucleotide)) %>% 
  mutate(Rowname = factor(Rowname, levels = unique(Rowname))) %>%
  filter(class != "Others")
head(data.df)


## row annotation
row_annotation_df <- data.df %>%
  select(TE, class) %>%
  distinct() %>% # keep the original arranged order
  as.data.frame() %>%
  mutate(class = factor(class, levels = TEclass_order) ) %>%
  arrange(class, TE) %>% # reorder
  tibble::column_to_rownames(var = "TE")
         
head(row_annotation_df)

row_annotation <- HeatmapAnnotation(df = row_annotation_df,
                                    col = list(class = TEclass_col),
                                    which = "row")

## build data matrix for plot
data.matrix <- data.df %>% 
  select(TE, Rowname, mean_freq) %>%
  spread(key = Rowname, value = mean_freq, fill = NA) %>%
  as.data.frame() %>%
  tibble::column_to_rownames(var = "TE")
# reorder by row_annotation
data.matrix <- data.matrix[rownames(row_annotation_df),]
write.csv(data.matrix, paste0("./11_views_TE_v2/07_1U10A_heatmap/HeatmapOf1U10A.mean_freq.csv"), quote = F, row.names = F)

  
### plot heatmap ###
col1 <- colorRamp2(c(0, 0.25, 1), c("#023e8a", "white", "#9e2a2b"))
column_split <- rep(1:(ncol(data.matrix) / 4), each = 4)

pdf(file=paste0("./11_views_TE_v2/07_1U10A_heatmap/HeatmapOf1U10A.pdf"),height=7,width=7)
### plot all ###
Heatmap(data.matrix, 
        column_title = "mean of nucleotide freq",
        left_annotation = row_annotation,
        cluster_columns = FALSE,
        cluster_rows = FALSE, row_dend_reorder = FALSE, # keep original order
        show_row_names = FALSE, show_column_names = TRUE,
        col = col1,
        column_split = column_split, # split by col
        name = "mean freq",
        use_raster = FALSE # keep original resolution
        )
  dev.off()
  

```

